Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-30 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: text on edgb and included missing cdf figs for eft

\subsection{Einstein-dilaton-Gauss-Bonnet}
\label{sec:results_edgb}

We start by presenting our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized posteriors of the coupling constant $\ell_{\rm GB}$,
for GW150914 (top panel) and GW200129 (middle panel), with and without the spin correction to the $(2,2,0)$
QNM. The bottom panel panel shows the joint posterior combining both events.
%
We see for both events events the $N_{\rm max} = 0$ posteriors are characterized by a peak away from zero.
%
This does not mean that we are observing a deviation from GR. We recall that
the deviations from GR in the ParSpec framework are controlled by $\gamma$,
which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{\rm s}}\right)^4\,.
\end{equation}
%
As shown in Fig.~xxx, $\gamma_{\rm GB}$ does indeed have a posterior with maximum support near zero, indicating consistency
between data and GR.
%
We also observe that the inclusion of the spin corrections (i.e., $N_{\rm max} = 1$) displaces the posteriors towards
smaller values of $\ell_{\rm GB}$.

As we stressed in Sec.~\ref{sec:remarks}, before we can draw any conclusions on
the allowed values for $\ell_{\rm GB}$ from our parameter estimation study, we
must first check whether the ``EFT''~\eqref{eq:eft_bound} and
``ParSpec''~\eqref{eq:parspec_bound} bounds are satisfied.
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the cumulative density function of the $\ell_{\rm GB}$ posteriors for GW150914 (GW200129), obtained
evaluating the integral~\eqref{eq:def_cdf} with $\Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_f$, solid lines), while varying $\varepsilon$.
%
For GW150914, we see that for the $N_{\rm max} = 0$ that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$, even at $\varepsilon =
1$. This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm GB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change when we include spin corrections to the EdGB QNM.
%
Altogether, \emph{we conclude that we cannot constrain EdGB gravity with our
present model}.
%
These conclusions are summarized in Table~\ref{tab:summary_edgb}.

@@

@@

>>>>

\subsection{Einstein-dilaton-Gauss-Bonnet gravity}
\label{sec:results_edgb}

We start by showing our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized posteriors of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel panel shows the joint posterior combining both events.
%
We see that for both events events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR from data.
%
Recall that the deviations from GR in the ParSpec framework are controlled
by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{\rm s}}\right)^4\,.
\end{equation}
%
As shown in Fig.~xxx, $\gamma_{\rm GB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between data
and GR.
%
We also observe that the inclusion of the spin corrections (i.e., curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm GB}$.

As we emphasized in Sec.~\ref{sec:remarks}, before drawing any conclusions on
the allowed values for $\ell_{\rm GB}$ from our parameter estimations, we
must first check whether the ``EFT''~\eqref{eq:eft_bound} and
``ParSpec''~\eqref{eq:parspec_bound} bounds are satisfied.
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the cumulative density function of the $\ell_{\rm GB}$ posteriors for GW150914 (GW200129), obtained
evaluating the integral~\eqref{eq:def_cdf} with $\Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_f$, solid lines), while varying $\varepsilon$.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$. This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm GB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we include spin corrections to the EdGB QNM.
%
Altogether, \emph{we conclude that we cannot constrain EdGB gravity with our
present model}.
%
These conclusions are summarized in Table~\ref{tab:summary_edgb}.

Let us contrast our results with those of Ref.~\cite{Lyu:2022gdr} which placed
the bound
%
$\ell_{\rm GB} \leqslant 1.33$~km
%
using the inspiral portion of the neutron star-black hole binaries GW200105 and GW200115~\cite{LIGOScientific:2021qlt} or
of Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, which placed the bound $\ell_{\rm GB} \leqslant 1.7$~km, by considering
a selection of black hole binaries from the GWTC-1 and GWTC-1 catalogs~\cite{}\footnote{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small, i.e., $\varphi \ll 1$, and take into consideration only the leading-order
scalar field contribution arising from the dilatonic coupling.}
%

% which relies on post-Newtonian calculations of the GW phase shift in this theory~\cite{Yagi:2011xp,Shiralilou:2020gah,Shiralilou:2021mfl}
% combined with parametrized post-Einsteinian (ppE) framework~\cite{Yunes:2009ke,Chatziioannou:2012rf,Tahura:2019dgr}.

Based on the later, but now here confirmed, within the assumptions of our waveform model,
in the case of comparable mass binaries.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_cdf_varying_threshold.pdf}
\caption{
}
\label{fig:cEFT_cdf}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_cdf_varying_threshold.pdf}
\caption{
}
\label{fig:qEFT_cdf}
\end{figure}
